Spherical harmonic analysis of faceted spheroids identifies shaping strategies and standardisation at Qianshangying (North China)

### 2025-08-22 analysis/paper/paper.qmd

To reduce the dimensionality and visualize the structure of shape variation, we applied a Uniform Manifold Approximation and Projection (UMAP) to the normalized power spectrum. The UMAP reveals distinct clusters in the first two dimensions of the shape space. These clusters correspond to reference models of sphere, rounded cube, ellipsoid, rounded box and disc-like (@fig-UMAP-combined). This demonstrates that these specimens occupy distinguishable regions in the shape space, with proximity to the sphere or elongation/flattening along specific axes emerging as key differentiating features. Most artifacts typologically classified as spheroids or sub-spheroids were grouped into the “sphere” , “rounded cube” and "ellipsoid" clusters, indicating their consistent, near-spherical shapes (@fig-UMAP-combined).

>>>>

To reduce the dimensionality and visualize the structure of shape variation, we applied a Uniform Manifold Approximation and Projection (UMAP) to the normalized power spectrum. The UMAP reveals distinct clusters in the first two dimensions of the shape space. These clusters correspond to reference models of sphere, rounded cube, ellipsoid, rounded box and disc-like (@fig-UMAP-combined). This demonstrates that these samples occupy distinguishable regions in the shape space, with proximity to the sphere or elongation or flattening along specific axes emerging as key differentiating features. Most artifacts typologically classified as spheroids or sub-spheroids were grouped into the “sphere” , “rounded cube” and "ellipsoid" clusters, indicating their consistent, near-spherical shapes (@fig-UMAP-combined).
